---
title: "Integrating the Effects of Climate Change and Invasive Species on Wetland Ecohydrology to Evaluate Management Options"
subtitle: ""
author: "Joe Shannon"
institute: "College of Forest Resources and Environmental Science,</br>Michigan Technological University"
date: "AGU Fall Meeting 2020"
output:
  xaringan::moon_reader:
    css: [xaringan-themer.css]
    self_contained: false
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(htmltools.dir.version = FALSE)
library(tidyverse)
library(lubridate)
library(jpshanno)
load_liberation_fonts()
```

```{r theming, include = FALSE}
library(xaringanthemer)
style_mono_light(base_color = "#1c5253",
                 background_color = "#FFF",
                 # footnote_font_size = "0.6em",
                 header_h1_font_size = "45px",
                 header_h2_font_size = "35px",
                 header_h3_font_size = "25px",
                 extra_css = list("li" = list("margin" = "0.75em 0")))
```
## Presentation Overview
Not just an outline of the presentation (Should be graphical: flowchart like
showing map of MI/Ash Importance, zoom in on wetlands, show three big drivers (EAB, Climate,
Mgmt), end with question about what it means). 
Present broad methodology overview 'I ran *statistical wetland
simulations* under current climate and future climate conditions to study the
interactions of invasive species, climate change, and forest management
decisions on forested wetland ecosystems'

---
## Problem Context
### Black Ash
.pull-left[
- Northern deciduous hardwood well adapted to wet/flood conditions  
- Dominant hardwood in northern forested wetlands
- Culturally important for indigienous communies
]
.pull-right[
**PICTURE OF BLACK ASH WETLAND**
**MAP OF Ash IV**
]

---
count:false
## Problem Context

### Black Ash & Emerald Ash Borer (EAB)

.pull-left[
- Invasive wood-borer
- First found in SE Michigan, USA in 2002
- Now in XXX states & provinces
- Impacts all North American ash
]
.pull-right[
**PICTURE OF EAB**
**FILL IN STATES IN ASH IV MAP SHOWING DETECTS**
]

---
count:false
## Problem Context

### Black Ash & EAB & Climate

.pull-left[
- Increased Precip Variability
- Decreased Summer Precip
- Decreased snowcover
]
.pull-right[
**figure from Byun or Nataro**
]

---
## Where Does this Leave Us?

**IMAGES OF FORESTED WETLAND, HERBACEOUS/SHRUB RESPONSE, & SEEDINGLINGS/PLANTING**

???
Talk about Iverson here?

---
## Data Collection
.pull-left[
- 14 black ash wetlands in Western Upper Peninsula of Michigan (map of Great Lakes)
- 2012 water level monitoring began
- Winter 2013-2014 simulated EAB infestation 
- Found decreased water level drawdown & increased late summer water levels
- Strong local/intermediate groundwater signal well into/through the growing season
]

.pull-right[
**Figures from sapflux & water level papers**
]

---
## Wetland Water Level Model Methodology
- Model daily water level change
- Statistical Model Based on Daily precip, minimum temperature, maximum temperature
- Model derivation flowchart:  
sy_mod  
intercept -> precip_rise -> gross_inflows  
PET : water availability on days without rain -> gross_outflows  
streamflow -> 

---
## Wetland Model Validation
- Might be able to do validations based on simulations from observed data
- 

---
## Climate Data Background
- GFDL-CM3  
- Figure from @byun-2018  
- LOCA downscaled daily projections  
- Future climate used years from 2070-2099  
- Selected 100 pixel time series from the study area to serve as 3000 annual 
weather sequences  

---
# Results under Current Climate

---
# Results under Future Climate
- Define wetland morphology groups

---
# Comparison of Results
- Compare within treatment between climates  
- Compare difference between treatments given climate  
- Any wet year vs dry year differences?
- Need to at least have basic met differences between simulations if any 
water level differences are present

---
## Future Work
- Bookend climate simulation selection - Current selection represents more
sensitive model and least control on emissions. Next will represent less
sensitive model  
- Improve Model through Rigorous Validation  
- Move from LOCA-projected daily values to spatially-explicit SWG with forcing
from LOCA-projections  
- Transition from discrete models to a distribution of possible parameters to 
capture a wider range of variation
- Compare my work with Iverson

???
- Improve from water_availability (need to get it from January 1st)
- Incorporate ARIMA components
